Rigidity of Expectation Formation: Additional Evidence from Desntify Forecasts

### 2019-06-26 paper/Draft.tex
IR sections continued. Almost finished the replication subsection.

%opening
\title{Testing Theories of Expectation Formation using Probabilistic Forecasts}
\author{Tao Wang}

@@

The theories on how different agents form expectations have proliferated over the past decade. On one hand, these theories built upon different micro foundations produce some similar macro patterns. For instance, the notion of information rigidity, or expectation rigidity could be micro founded by many different theories, but mostly generate sluggish response to new information.

On the other hand, however, there are important and subtle differences in testable predictions from these theories for both individual forecasts and aggregate moments of the forecasts. My goal of this exercise is to derive the testable predictions based on various theories, and utilize the probablistic questions from both professional forecasters survey and household survey to test them.

@@

We follow \cite{Manski} to estiamte the density distribution of each individual surveyee for SPF. New York Fed has followed the same procedures for SCE and directly provided their estimate of uncertainty. Therefore, I directly use them.

@@

@@

Another widely spoken fact is that uncertainty is positively correlated with the level of the inflation. We look this hypothesis by ploting uncertainty against the realized inflation.

@@

\begin{figure}[h]\label{RevisionHist}
	\centering
	\includegraphics[width=6cm]{figures/PRCCPIMean_rv_hist.png} 
	\includegraphics[width=6cm]{figures/PRCCPIVar_rv_hist.png}  \\
	\smallskip
	\includegraphics[width=6cm]{figures/PRCPCEMean_rv_hist.png} 
	\includegraphics[width=6cm]{figures/PRCPCEVar_rv_hist.png}  \\
	\smallskip
		\includegraphics[width=6cm]{figures/SCEMean_rv_hist.png} 
	\includegraphics[width=6cm]{figures/SCEVar_rv_hist.png}  \\
	\caption{Distribution of Revision in Forecasts}
\end{figure}

@@

 \cite{Coibion2012can} explores the implications of expectation rigidity models using externally identified shocks to inflation.  Unlike the time series regressions that test null hypotheses of FIRE in previous section, this approach allows a more straightforward comparison between the emprical evidence from surveys and the theoriecal predictions illustrated in Figure \ref{ir_pop}. The difficulty of this approach lies in identifying a shock $\omega_t$ that is a pure innovation to inflation at time $t$.

In this paper I use three types of shocks: technology shock\cite{gali1999technology}, oil price shock\cite{hamilton1996happened} and monetary policy shocks \cite{laseen2011anticipated}. The first two are the same in \cite{Coibion2012can}. \footnote{ Another shock in \cite{Coibon2012can} is the news shock by \cite{barsky2011news}. This will be included in later version of the paper }

(1) Technology shock is identified from a three variate structural vector autoregression(SVAR) model  (output per capita, hours and inflation) by imposing a widely used long-run restriction that non-technology shocks have no long-run output impacts. \footnote{An recent alternative to long-run restriction approach is the ''max-share appraoch'' developed by \cite{francis2014flexible}. It is shown to have better small-sample properties. In particular,  the approach defines technology shock as the variation of current period labor productivity (output per hour) that explains the maximum fraction of future labor productivity over an exogenously determiend horizon. The resutls using this shock will be included in a later version of the paper. For the replication purpose now, I stay with \cite{gali1999technology}}. The sample period used to identify the shock is 1949-2019, while the period where it enters my analysis is the same to the period of survey data: 2007-2019 for SPF and 2013-2019 for SCE.

@@

In Figure \ref{ReplicateCoibionBefore2007}, I replicate the results from  \cite{coibion2012can}. The period is between 1987-2007.  In the right column, I do the same exercise with monetary policy shocks.

>>>>

%opening
\title{Rigidity of Expectation Formation: Additional Evidence from Desntify Forecasts}
\author{Tao Wang}

@@

The theories on how agents form expectations in ways deviating from rational expectation(RE) benchmark have proliferated over the past decade. On one hand, these theories built upon different micro foundations produce some similar macro patterns. For instance, the notion of information rigidity, or expectation rigidity could be micro founded by many different theories, but mostly generate sluggish response to new information.  On the other hand, there are important and subtle differences in testable predictions from these theories for both individual forecasts and aggregate moments of the forecasts. This paper extends the current literature in this front by deriving the testable predictions from various theories on uncertainty and empirically test them utilizing the previously ignored probablistic forecast from both professional forecasters survey and household survey. Overall, the paper finds additional evidence from uncertainty consistent with the rigidity models.

@@

We follow \cite{engelberg2009comparing} to estiamte the density distribution of each individual surveyee for SPF. New York Fed has followed the same procedures for SCE and directly provided their estimate of uncertainty. Therefore, I directly use them.

@@

Throughout the paper, I use two measures of Inflation, headline CPI inflation and PCE inflation. The forecasts-based measures are computed and used correspondingly.

@@

Another widely made statement is that uncertainty is positively correlated with the level of the inflation. I look into this hypothesis by ploting uncertainty against the realized inflation.

@@

\begin{figure}[h]\label{RevisionHist}
	\centering
	\includegraphics[width=6cm]{figures/PRCCPIMean_rv_hist.png} 
	\includegraphics[width=6cm]{figures/PRCCPIVar_rv_hist.png}  \\
	\smallskip
	\includegraphics[width=6cm]{figures/PRCPCEMean_rv_hist.png} 
	\includegraphics[width=6cm]{figures/PRCPCEVar_rv_hist.png}  \\
	\caption{Distribution of Revision in Forecasts}
\end{figure}

@@

 \cite{coibion2012can} explores the implications of expectation rigidity models using externally identified shocks to inflation.  Unlike the time series regressions that test null hypotheses of FIRE in previous section, this approach allows a more straightforward comparison between the emprical evidence from surveys and the theoriecal predictions illustrated in Figure \ref{ir_pop}. The difficulty of this approach lies in identifying a shock $\omega_t$ that is a pure innovation to inflation at time $t$.

In this paper I use three types of shocks: technology shock\cite{gali1999technology}, oil price shock\cite{hamilton1996happened} and monetary policy shocks \cite{laseen2011anticipated}. The first two are the same in \cite{coibion2012can}. \footnote{ Another shock in \cite{coibon2012can} is the news shock by \cite{barsky2011news}. This will be included in later version of the paper.}

(1) Technology shock is identified from a three variate structural vector autoregression(SVAR) model  (output per capita, hours and inflation) by imposing a widely used long-run restriction that non-technology shocks have no long-run output impacts. Based on Bayes information criteria, I use the lag of 4 quarters as in \cite{coibion2012can} \footnote{A recent alternative to long-run restriction approach is the ``max-share appraoch'' developed by \cite{francis2014flexible}. It is shown to have better small-sample properties. In particular,  the approach defines technology shock as the variation of current period labor productivity (output per hour) that explains the maximum fraction of future labor productivity over an exogenously determiend horizon. The resutls using this shock will be included in a later version of the paper. For the replication purpose now, I stay with \cite{gali1999technology}}. The sample period used to identify the shock is 1949-2019, while the period where it enters my analysis is the same to the period of survey data: 2007-2019 for SPF and 2013-2019 for SCE.

@@

For a direct comparison of my results and \cite{coibion2012can}, I first explore the impulse responses along one-standard-deviation confidence band of inflation $y_{t+k}$, and SPF forecast errors $\overline {FE}_{t+h|t+k}$ to one standard deviation of each shock at time $t$, and SPF disagreements $\overline Disg_{t+h|t+k}$ and average uncertainty $\bar\sigma^2_{t+h|t+k}$ to the absolute value of shocks. The sample period is the same to that in Figure 2 from \cite{coibion2012can}: 1987-2007. The only difference is that CPI and PCE inflation are used instead of GDP deflator as in their paper.

The underlying regression that produces impulse response estimates is the Equation \ref{IRRegression}, according to which inflation(or moments) in current quarter is a function of 4-period lagged of itself  and the current and previous period realized shocks. The estimate of autoregression coefficient $\beta_\tau$ and shock-specific coefficient $\beta_{S,k}$ for diffrent lags can be used to compute impulse response of the inflation as well as forecasting moments.

\begin{eqnarray}\label{IRRegression}
y_{t}  = \alpha + \sum^4_{\tau=1} \beta_{\tau} y_{t-\tau} +  \sum^1_{k=0} \sum_{S \in \{O,T\}}\beta_{S,k}\omega_{S,t-k} + \epsilon_t 
\end{eqnarray}

FIRE predicts $\beta_{S,k}=0\quad \forall k=1$ for the regressions with forecasts moments as dependent variable. i.e. forecasts react only to contemporaneous shocks instataneously therefore.

Standard errors are computed using the parametric bootstrapping approach as in \cite{coibion2012can} paper.

Responses to productivity and oil price shocks presented in the left panel of the Figure \ref{ReplicateCoibionBefore2007} are broadly consistent with \cite{coibion2012can} despite a wider confidence band possibly due to the change in different measures. First, one standard deviation of productivity unsurprisingly brings down headline inflation by around $10 \%$  in the first quarter and the negative impacts gradually mute since then till the 7-th quarter. Oil price shock of one standard deviation increases headline inflation by around $30\%$ in the first quarter and the inflationary impacts last for 6 quarters. PCE inflation responds to both the shocks in a similar manner with slightly different magnitudes. The responses coefficients are statistically significant as the $95\%$ confidence intervel stays on one side of zero for both shock for the same period.

Second, population forecast errors respond to shocks in a sluggish manner as predicted in rigidity models. This is illustrated using both CPI and PCE forecasts in the second graph of the left panel in Figure \ref{ReplicateCoibionBefore2007}. As we define the forecast error as forecast minus realized value, forecast error rises(drops) after a positive(negative) shock and gradully return to zero around 4-5th quarters after the shock.

Both SE and NI models predict disagreements rise after a shock and gradully converge to zero over the long-run.
